---
title: "Recirculation Speed"
author:
    - name: "Jem Arnold "
    # - name: "Jem Arnold [{{< fa brands github >}}](https://github.com/jemarnold) [{{< fa brands bluesky >}}](https://bsky.app/profile/jemarnold.bsky.social) [{{< ai researchgate >}}](https://researchgate.net/profile/Jem-Arnold) [{{< ai orcid >}}](https://orcid.org/0000-0003-3908-9447)"
date: last-modified
format:
    html:
        embed-resources: true
        toc: false
    gfm:
        output-file: README.md
knitr:
    opts_chunk:
        collapse: true
        comment: '#>'
        dpi: 300
        fig.asp: 0.707
        out.width: "100%"
        fig.align: "center"
        fig.path: "figures/"
---

```{r setup}
#| code-fold: true
#| code-summary: "Show setup code"

library(tidyverse)
library(JAPackage)

df <- readRDS(file = r"(C:\R-Projects\data-vis-threads\2026-06-25\recirculation-df)")

points_df <- tibble::tribble(
    ~time , ~VL_ICG_filt ,
     8.62 , 15.34        ,
    18.82 , 12.14        ,
    29.66 , 11.4         ,
)

theme_set(theme_JA(plot.title.position = "plot"))
```


How long does it take for blood to make a full round-trip throughout the entire body during exercise?

Around 10-sec!

That's way faster than I intuitively would have expected! 

But we're only reproducing what's been known since at least the 1950s 👇



```{r}

ggplot(df, aes(x = time)) +
        labs(
            title = str_glue(
                "<span style = 'color:green4'>**NIRS-ICG**</span>",
                " dye tracer monitored at ",
                "***Vastus Lateralis*** ",
                "during cycling intervals"
            ),
            subtitle = "Systemic blood recirculation time ≈ 10 seconds"
        ) +
        scale_x_continuous(
            name = "Time (mm:ss)",
            breaks = seq(0, 60, 15),
            labels = mnirs::format_hmmss,
            expand = expansion(0.02)
        ) +
        scale_y_continuous(
            name = "NIRS-ICG Amplitude (μM)",
            expand = expansion(0.01)
        ) +
        scale_colour_manual(
            aesthetics = c("colour", "fill"),
            values = c(VL = "green4"),
        ) +
        geom_vline(xintercept = 0, linetype = "dotted") +
        geom_line(aes(y = VL_ICG, colour = "VL"), alpha = 0.3) +
        geom_line(aes(y = VL_ICG_filt, colour = "VL")) +
        geom_point(
            data = points_df,
            aes(x = time, y = VL_ICG_filt, colour = "VL"),
            size = 3, shape = 21, stroke = 1, fill = "white"
        ) + 
        annotate(
            "text", x = 0, y = 4, label = "Dye injection",
            angle = 90, vjust = -0.5, hjust = 0.5, colour = "green4"
        ) +
        # one arrow from text label to first peak, then bouncing peak-to-peak
        annotate(
            "curve",
            x = c(45, head(points_df$time, -1) + c(1, 0.5)),
            y = c(5, head(points_df$VL_ICG_filt, -1) + c(0, 1)),
            xend = points_df$time + c(0, -0.5, -0.5),
            yend = points_df$VL_ICG_filt + c(-1, 1, 1),
            arrow = arrow(length = unit(3, "mm"), type = "closed"),
            linewidth = 0.8, colour = "grey20", curvature = -0.4,
        ) +
        annotate(
            "label", x = 40, y = 4,
            label = "Dye bolus concentration peaks\nunder the muscle NIRS probe.\nRecirculation every ~10-sec", fill = "white", size = 5, hjust = 0.5, vjust = 0.5
        )
```

A figure showing near-infrared spectroscopy (NIRS) tracing over the vastus lateralis quadricep muscle during a cycling VO2max interval, with indocyanine green (ICG) dye tracer injected at time 0:00; as the dye appears with blood flow under the probe, the optical signal rapidly reaches a peak at ~10-seconds after first injection. The concentration decays briefly, then rises to a second and third peak at ~10-sec increments, showing transit time around the entire systemic circulation (from quadriceps, via venous blood back through lungs to heart, and out again in arterial circulation to the quadriceps). Figure title: "NIRS-ICG dye tracer monitored at vastus lateralis during cycling intervals. Systemic blood recirculation time ≈ 10 seconds". Text label: "Dye bolus concentration peaks under the NIRS probe. Recirculation every ~10-sec".