---
title: Ultrasound Adipose Tissue Thickness ~ Calliper Skinfold Correlation
author:
    name: Jem Arnold [{{< fa brands github >}}](https://github.com/jemarnold/mnirs){target=\"_blank\"} [{{< fa brands linkedin >}}](https://www.linkedin.com/in/jem--arnold/){target=\"_blank\"} [{{< ai researchgate >}}](https://researchgate.net/profile/Jem-Arnold){target=\"_blank\"} [{{< ai orcid >}}](https://orcid.org/0000-0003-3908-9447){target=\"_blank\"}
    email: jem@jemarnold.ca
    affiliation:
        name: |
            School of Kinesiology\
            The University of British Columbia\
            Vancouver, BC, Canada
date: last-modified
format:
    html:
        self-contained: true
        toc: true
        code-fold: true
    gfm:
        output-file: README.md
execute:
    echo: true
    message: false
    warning: false
    dpi: 300
    fig-width: 8.66
    fig-asp: 0.5
    fig-align: center
    out-width: "100%"
knitr:
    opts_chunk:
        fig.path: "figures/"
---

```{r}
#| label: setup
#| code-summary: "Show setup code"

library(tidyverse)
library(ggpmisc)
library(mnirs)
library(lme4)
library(lmerTest)
library(rmcorr)
library(emmeans)
library(marginaleffects)

theme_set(JAPackage::theme_JA(base_size = 16, base_family = "sans"))
```

::: {.callout-note}
## Methods

* Sample is **32** young, well-trained female and male competitive cyclists in the Netherlands in 2022.

* Mix of patients with blood flow limitations in the iliac artery and healthy participants (pooled for this analysis).

* Ultrasound adipose tissue thickness (ATT) vs calliper skinfold thickness (SKF).

* Measurements taken at bilateral quadriceps (VL) at 1/3 distance from patella to greater trochanter.

* Skinfold values divided by 2, comparable to ATT.

:::

## Data

```{r}
#| echo: false

df <- readRDS("att-skf-df.rds")
```

```{r}
#| label: data

df

## participants by sex
df |>
    distinct(id, sex) |>
    count(sex)
```

## Figure 1: ATT vs Skinfold correlation

* Dotted line: line of identity.

* Dashed grey line: full group regression.

* Coloured lines: blue is males, red is females.

```{r}
#| label: correlation-plot
#| code-summary: "Show figure code"
#| fig-asp: 0.667

# fmt: skip
ggplot(df, aes(x = att, y = skf, colour = sex, fill = sex)) +
    labs(
        title = "Correlation between Ultrasound **Adipose Tissue**<br>and Calliper **Skinfold** measurements",
        subtitle = str_glue("n = 32 (9 F, 23 M) well-trained cyclists, bilateral quadriceps (VL)"),
        x = "Adipose Tissue Thickness (mm)",
        y = "Skinfold Thickness ÷ 2 (mm)"
    ) +
    scale_colour_mnirs() +
    scale_fill_mnirs() +
    coord_cartesian(xlim = c(0, NA), ylim = c(0, NA)) +
    scale_x_continuous(
        breaks = seq(0, 12, 2),
        expand = expansion(c(0, 0.03))
    ) +
    scale_y_continuous(
        breaks = seq(0, 18, 2),
        expand = expansion(c(0, 0.03))
    ) +
    geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
    geom_point(fill = "white", size = 4.4, shape = 21, stroke = 1.5) +
    geom_point(colour = "white", size = 4.2) +
    geom_point(size = 5.0, shape = 21, stroke = 0, alpha = 0.6) +
    geom_smooth(formula = y ~ x, method = "lm", se = TRUE, alpha = 0.2) +
    geom_smooth(
        inherit.aes = FALSE,
        aes(x = att, y = skf),
        formula = y ~ x, method = "lm", se = TRUE,
        colour = "grey10", linetype = 2, alpha = 0.4
    ) +
    ## full group stats
    stat_poly_eq(
        use_label(c("eq"), other.mapping = aes(x = att, y = skf)),
        inherit.aes = FALSE,
        formula = y ~ x, method = "lm",
        label.x = c(0.03), label.y = c(0.97),
        colour = "grey10", size = 4.5
    ) +
    stat_correlation(
        use_label(c("r", "rr", "p"), other.mapping = aes(x = att, y = skf)),
        inherit.aes = FALSE,
        method = "pearson", small.r = FALSE, small.p = TRUE,
        label.x = c(0.03), label.y = c(0.92),
        colour = "grey10", size = 4.5
    ) +
    ## per-sex stats
    stat_poly_eq(
        use_label(c("eq")),
        formula = y ~ x, method = "lm",
        label.x = c(0.03), label.y = c(0.83, 0.69), size = 4.5
    ) +
    stat_correlation(
        use_label(c("r", "rr", "p")),
        method = "pearson", small.r = FALSE, small.p = TRUE,
        label.x = c(0.03), label.y = c(0.78, 0.64), size = 4.5
    )
```


## Repeated-measures correlation

* Accounts for bilateral (L & R) measurements within participant.

```{r}
#| label: rmcorr

rmcorr(
    participant = id,
    measure1 = att,
    measure2 = skf,
    dataset = df,
)
```


## Figure 2: Skinfold bias & equivalence by sex

* Bias (Δ) = SKF ÷ 2 − ATT, per leg. Positive values: skinfold overestimates adipose tissue.

* Individual points & estimated marginal means [90% CI] from mixed effects model `delta ~ sex * leg + (1 | id)`, pooled over legs.

* Two one-sided tests (TOST) with equivalence bounds chosen at **±1 mm** (shaded region): "*is SKF ÷ 2 within ±1 mm of ATT on average?*" Equivalence is supported when the full 90% CI lies within the shaded region.

* Suggests a sex interaction, where skinfolds tend to overestimate adipose tissue more in females than in males.

* Weak/inconclusive equivalence, sensitive to a couple of outliers in this small dataset.

```{r}
#| label: equivalence
#| code-summary: "Show code"

delta_df <- mutate(df, delta = skf - att)

delta_fit <- lmer(delta ~ sex * leg + (1 | id), data = delta_df)

## mean bias by sex, averaged over legs
delta_emm <- emmeans(delta_fit, ~sex)

## female vs male difference in bias
delta_contrast <- contrast(delta_emm, "revpairwise") |>
    confint() |>
    as.data.frame()

delta_contrast

## equivalence bound (mm)
eq_bound <- 1

## TOST p-values with matching 90% CI, by sex, averaged over legs
## population-level predictions (re.form = NA), between-participant df
eq_df <- marginaleffects::avg_predictions(
    delta_fit,
    by = "sex",
    re.form = NA,
    conf_level = 0.90,
    equivalence = c(-eq_bound, eq_bound),
    df = n_distinct(delta_df$id) - 2
) |>
    as.data.frame()

eq_df
```

```{r}
#| label: equivalence-plot
#| code-summary: "Show figure code"
#| fig-asp: 0.667
# fmt: skip
ggplot(delta_df, aes(x = delta, y = sex, colour = sex, fill = sex)) +
    labs(
        title = "Calliper **Skinfold** bias relative to Ultrasound **Adipose Tissue** by sex",
        subtitle = str_glue_data(
            delta_contrast,
            "Estimate [90% CI], equivalence bounds **±{eq_bound} mm**<br>",
            "*Female − Male* bias: {round(estimate, 1)} [95% CI {round(lower.CL, 1)}, {round(upper.CL, 1)}] mm"
        ),
        x = "Skinfold ÷ 2 − Adipose Tissue (mm)",
        y = NULL
    ) +
    theme(
        legend.position = "none",
        plot.title = ggtext::element_textbox(size = 18),
        axis.text.y = element_text(size = 16, colour = "black", face = "bold"),
    ) +
    scale_colour_mnirs() +
    scale_fill_mnirs() +
    scale_x_continuous(breaks = seq(-2, 8, 1)) +
    ## male on top
    scale_y_discrete(limits = rev, expand = expansion(add = c(0.3, 0.7))) +
    annotate(
        "rect", xmin = -eq_bound, xmax = eq_bound, ymin = -Inf, ymax = Inf,
        fill = "palegreen3", alpha = 0.5
    ) +
    geom_vline(xintercept = c(-eq_bound, eq_bound), linetype = "dashed") +
    geom_vline(xintercept = 0) +
    ## narrower spread for smaller female group
    map2(c("Male", "Female"), c(0.25, 0.1), \(.sex, .width) {
        .df <- filter(delta_df, sex == .sex)
        list(
            ggbeeswarm::geom_quasirandom(
                data = .df, width = .width,
                fill = "white", size = 4.4, shape = 21, stroke = 1.5,
            ),
            ggbeeswarm::geom_quasirandom(
                data = .df, width = .width,
                colour = "white", size = 4.2,
            ),
            ggbeeswarm::geom_quasirandom(
                data = .df, width = .width,
                size = 5.0, shape = 21, stroke = 0, alpha = 0.6,
            )
        )
    }) +
    geom_pointrange(
        data = eq_df,
        aes(x = estimate, xmin = conf.low, xmax = conf.high),
        position = position_nudge(y = 0.35),
        size = 2, linewidth = 1.5, shape = 21, fill = "white", stroke = 2
    ) +
    geom_text(
        data = eq_df,
        aes(x = estimate, label = str_glue("TOST p = {round(p.value.equiv, 2)}")),
        position = position_nudge(y = 0.55),
        size = 5, fontface = "bold"
    )

```


```{r}
#| label: session-info
#| code-summary: "Show sessionInfo()"
#| collapse: true

sessionInfo()
```
