---
title: Comparison of Moxy and Train.Red
author:
    name: Jem Arnold [{{< fa brands github >}}](https://github.com/jemarnold/mnirs){target=\"_blank\"} [{{< fa brands linkedin >}}](https://www.linkedin.com/in/jem--arnold/){target=\"_blank\"} [{{< ai researchgate >}}](https://researchgate.net/profile/Jem-Arnold){target=\"_blank\"} [{{< ai orcid >}}](https://orcid.org/0000-0003-3908-9447){target=\"_blank\"}
    email: jem@jemarnold.ca
    affiliation:
        name: |
            School of Kinesiology\
            The University of British Columbia\
            Vancouver, BC, Canada
date: last-modified
format:
    html:
        self-contained: true
        toc: true
        code-fold: true
    gfm:
        output-file: README.md
execute:
    echo: true
    message: false
    warning: false
    dpi: 300
    fig-width: 7
    fig-asp: 0.5
    fig-align: center
    out-width: "100%"
knitr:
    opts_chunk:
        fig.path: "figures/"
---

When I've simultaneously tested **Moxy** and **Train.Red** previously, I've been satisfied that they mostly follow the same kinetics, when rescaled to an equivalent range.

::: {.callout-note}
## Methods

* **Moxy 5** running firmware *1.6.10* recording at **10 Hz**, inter-optode spacing **25mm**.

* **Train.Red FYER** running firmware *2.0.34* recording at **10 Hz**, inter-optode spacing **35mm**.

* Moxy positioned on the **proximal right VL** at around 1/2 distance. Train.Red positioned on the **distal right VL** around 1/3 distance from patella. Around 5 cm between sensors.

* Calliper skinfolds: distal = 5 mm, proximal = 6 mm.

* Both recording to *PerfPro* app via Ant+, receiving at **2 Hz**.

* Train.Red had a few inconsequential signal dropouts, so I'm removing them as outliers.

* Train.Red is smoothed onboard with unspecified parameters. Moxy is raw signal. So I'm filtering Moxy only with a Butterworth low-pass filter, order = 2, frequency cutoff = 0.15 Hz (6.67 cycles/sec).

:::

::: {.callout-note}
## Protocol

* 3-min seated rest baseline.

* 8-min resting ischaemic occlusion.

* 5-min recovery hyperaemia.

* 20-min moderate treadmill running, purpose to reach "warmed-up" working muscle vasodilation.

* 5-min resting exercise recovery.

* 8-min ischaemic occlusion.

* 2-min recovery hyperaemia (then the Train.Red sensor battery died).

Dashed lines in the plot below

:::

```{r}
#| label: setup
#| code-summary: "Show setup code"

library(dplyr)
library(ggplot2)
library(mnirs)

theme_set(theme_mnirs(base_size = 16))
```

```{r}
#| label: files
#| echo: false

files <- list.files(
    r"(C:\OneDrive - UBC\JAData\)",
    pattern = "moxy-train.red.*.xlsx",
    full.names = TRUE
)
```

```{r}
#| label: read-file-one
#| code-summary: "Show figure code"

session_one <- read_mnirs(
    files[1],
    nirs_channels = c(
        train.red_distal = "smo2_right_distal",
        moxy_proximal = "smo2_right_proximal"
    ),
    time_channel = c(time = "Time"),
) |>
    replace_mnirs(
        nirs_channels = train.red_distal,
        outlier_cutoff = 3,
        span = 15
    ) |>
    tail(-10) |> 
    head(-120) |>
    filter_mnirs(
        nirs_channels = moxy_proximal,
        method = "butterworth",
        order = 2,
        fc = 0.15
    ) |>
    create_mnirs_data(nirs_channels = c(train.red_distal, moxy_proximal))

plot(session_one, time_labels = TRUE) +
    coord_cartesian(ylim = c(0, 100)) + 
    geom_vline(xintercept = c(3, 11, 16, 36, 41, 49, 52)*60, linetype = 2)
```

## Rescale amplitudes

Test one: rescale both signals to their ischaemic calibration range.

```{r}
#| label: rescale
#| code-summary: "Show figure code"

session_one |>
    rescale_mnirs(
        group_channels = "distinct",
        range = c(0, 100)
    ) |> 
    plot(time_labels = TRUE)
```

## The first half

The first half of the session have equivalent amplitude and kinetics: the first occlusion, hyperaemic recovery, through the first half of the exercise interval are essentially overlapping.

The signals begin to diverge around half way through exercise. This could be attributed to the different sensor positions and expected differences in perfusion. Although I would have expected the distal Train.Red to show greater deoxygenation.

Alternatively, Train.Red could be treating the tissue signal homogenously, and could be picking up more of the expected cutaneous thermoregulatory perfusion. While Moxy attempts to prioritise the deeper signal and returns more deoxygenation from the muscle.

## The second occlusion

The second occlusion in the "warmed-up" state after exercise is more different. Moxy reached zero very quickly and essentially lost signal. We have seen this before in Moxy typically in highly vascularised, lean males, and speculate that the signal could be near-completely absorbed by blood volume before returning to the receiver.

## Isolate the second occlusion

Rescaling to the second occlusion only reveals the kinetics are more in line, especially during reoxygenation.

I interpret the noise in Train.Red at the bottom of the occlusion to be basically equivalent to the 'zero' signal in Moxy. They both represent weak/lost signal.

Train.Red continues to deoxygenate for approximately 60-sec longer than Moxy before bottoming out.


```{r}
#| label: second-half
#| code-summary: "Show figure code"

session_one |>
    tail(-3000) |>
    rescale_mnirs(
        group_channels = "distinct",
        range = c(0, 100)
    ) |>
    plot(time_labels = TRUE)
```

# Repeating the trial

::: {.callout-note}
## Methods

* As previous, with **Moxy** on **distal right VL** and **Train.Red** on **proximal right VL**.

* Train.Red outliers removed and Moxy filtered as previously.

:::

::: {.callout-note}
## Protocol

* 3-min seated rest baseline.

* 6.5-min resting ischaemic occlusion. Until visual Train.Red plateau reached.

* 4-min recovery hyperaemia.

* 20-min moderate cycling as warm-up stimulus this time.

* 6-min resting exercise recovery.

* 5-min ischaemic occlusion. Until visual Train.Red plateau reached.

* 4-min recovery hyperaemia.

Dashed lines in the plot below

:::

```{r}
#| label: read-file-two
#| code-summary: "Show figure code"

session_two <- read_mnirs(
    files[2],
    nirs_channels = c(
        moxy_distal = "smo2_right_distal",
        train.red_proximal = "smo2_right_proximal"
    ),
    time_channel = c(time = "Time"),
) |>
    replace_mnirs(
        nirs_channels = train.red_proximal,
        outlier_cutoff = 3,
        span = 15
    ) |>
    mutate(
        train.red_proximal = replace(
            train.red_proximal,
            between(time, 1056.88, 1061.29),
            NA
        )
    ) |>
    tail(-10) |>
    filter_mnirs(
        nirs_channels = moxy_distal,
        method = "butterworth",
        order = 2,
        W = 0.12
    ) |>
    create_mnirs_data(nirs_channels = c(train.red_proximal, moxy_distal))

plot(session_two, time_labels = TRUE) +
    coord_cartesian(ylim = c(0, 100)) +
    geom_vline(xintercept = c(3, 9.5, 13.5, 36, 42, 47, 51) * 60, linetype = 2)
```

## Rescale

```{r}
#| label: rescale-two
#| code-summary: "Show figure code"

session_two |>
    rescale_mnirs(
        group_channels = "distinct",
        range = c(0, 100)
    ) |>
    plot(time_labels = TRUE)
```

I was surprised at the Moxy dropouts during both occlusions, making it more difficult to compare kinetics. It may be related to the same issue as speculated above, over the distal VL, with less adipose tissue and expecting greater deoxygenation.

The rest of the file agrees quite well after rescaling.

## Isolate the second occlusion

```{r}
#| label: second-half-two
#| code-summary: "Show figure code"

session_two |>
    tail(-3000) |>
    rescale_mnirs(
        group_channels = "distinct",
        range = c(0, 100)
    ) |>
    plot(time_labels = TRUE)
```

This session leaves me feeling the Moxy is missing real physiological range that Train.Red is picking up. Which may be compounded in this case by an interaction of (A) the distal vs proximal muscle locations, (B) the sensor optode widths/depths, and (C) the tissue homogeneity signal processing philosophies.

## Comparing kinetics

Without rescaling amplitudes we can compare reoxygenation kinetics from the two occlusions and end of exercise in each session.


::: {.callout-note}
## Two-phase exponential-linear kinetics

Compare `tau`; the time constant in seconds, and `MRT`; the mean response time of the time delay (`TD`) plus `tau`, within each of the three intervals (occlusion, exercise, occlusion).

`texc` is the excursion time where the fast exponential kinetics transition into the linear drift.

They are mostly within a couple of seconds. Exercise reoxygenation is the most different. Session one shows closer values than Session two. Purely qualitative comparisons here.

`Slope_B` is the linear asymptote slope, and is expected to be different related to the device amplitudes.

:::



### Session one

```{r}
#| label: plot-kinetics-one
#| code-summary: "Show figure code"

extract_intervals(
    session_one,
    start = by_time(11 * 60, 36 * 60, 49 * 60),
    span = c(-30, 180)
) |>
    analyse_kinetics(
        method = "exponential-linear"
    ) |> 
    print() |> 
    plot(label_size = 3)
```

### Session two

```{r}
#| label: kinetics-two
#| code-summary: "Show figure code"

extract_intervals(
    session_two,
    start = by_time(9.5*60, 36*60, 47*60),
    span = c(-30, 150)
) |> 
    analyse_kinetics(
        method = "exponential-linear"
    ) |> 
    print() |> 
    plot()
```

```{r}
#| label: session-info
#| code-summary: "Show sessionInfo()"
#| collapse: true

sessionInfo()
```
